Contextual Combinatorial Cascading Bandits

### 2015-12-10 contextual combinatorial cascading bandits.tex

To state our main theorem, we need some definitions. Let
$$
f_{\wedge}^* = \min_{1 \leq t \leq T} f_{\wedge, t}^{\ast}
$$
denote the best expected reward of all time under the conjunctive objective.
Let $p_{t, A}$ (resp., $p_{\wedge, t, A}$) be the {\em probability of full
	observation of $A$}, that is the probability that all base arms of $A = (a_1, \ldots, a_{\abs{A}})$ are observed under the disjunctive objective (resp., conjunctive objective) when under the context of time $t$, that is 
\begin{align*}
&p_{t, A} = \prod_{k=1}^{\abs{A}-1} (1 - \theta_{\ast}^{\top} x_{t, a_k^t}),\\
&p_{\wedge, t, A} = \prod_{k=1}^{\abs{A}-1} (\theta_{\ast}^{\top} x_{t, a_k^t}).
\end{align*}
We also define $p^* = \min_{1 \leq t \leq T} \min_{A \in \cS} ~ p_{t, A}$ to be the minimal probability over all time that an action could have all base arms observed under the disjunctive objective. The following is the main theorem on the regret achieved by our C$^3$-UCB algorithm, for both the disjunctive and conjunctive objectives.
\begin{theorem}
\label{thm:main}
Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_{K} \geq 0$ and 
$\gamma_k' = 1 - \gamma_k, \gamma'_K \leq \frac{1}{4}f_{\wedge}^{\ast}$.
Let regularization parameter $\lambda$ satisfy $\lambda \geq C_\gamma$, where $C_\gamma = \sum_{k=1}^{K} \gamma_k^2$. Then for any $\delta > 0$, with probability at least $1 - \delta$, the cumulative expected regrets of our algorithm, C$^3$-UCB for both the conjunctive and disjunctive objectives, satisfy
\begin{align}
R(T) &\le \frac{\sqrt{2}}{p^*} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))}  \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2}\right) } + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{p^*} \sqrt{TK} \ln T\right),
\end{align}
\begin{align}
R_{\wedge}(T) &\le \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2}\right) } + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{f_{\wedge}^{\ast}}\sqrt{TK} \ln T\right).
\end{align}
%where
%\begin{align*}
%C_\gamma = \sum_{k=1}^{K} \gamma_k^2 = \sum_{k=1}^{K} (1 - \gamma_k')^2 \in [1, K].
%\end{align*}
\end{theorem}

@@

\begin{lemma}
\label{lem:DeltaEsimateWithP*}
Let $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$ and $\gamma_k' = 1 - \gamma_k$. 
For any time $t$ and the chosen action $\bA_t = (\ba_1^t, \ldots, \ba_{\abs{A}}^t )$, we have $f(A_t^*, \bU_t) \leq f(A, \bU_t)$. 
Suppose the equation (\ref{eq:estimateTheta}) holds for time $t-1$. Then
$$
\EE_t[R(t, \bA_t)] \leq \frac{1}{p^*} \EE_t \left[ \bbeta_{t-1}(\delta) \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
$$
If we have
% $$
% \gamma_K \geq (1+\frac{1}{n_0})f^{\ast}, \qquad 
% $$
$$
\gamma_K' \le \frac{1}{4} f_{\wedge}^{\ast} = \frac{1}{4} \min_{1 \leq t \leq T} f_{\wedge, t}^{\ast},
$$
then
\begin{align*}
% &E[\Delta_{t, \bA_t}|\cH_t] \leq (n_0+1) \bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}},\\
&\EE_t [R_{\wedge}(t, \bA_t) ] \leq \frac{8}{f_{\wedge}^{\ast}} \EE_t \left[ \bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
\end{align*}
\end{lemma}
\begin{proof}
For the disjunctive case,  
\begin{align}
&\EE[R(t, \bA_t)|\cH_t]  \nonumber \\
=& \EE \left[ \left. R(t, \bA_t) \EE \left[ \left. \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t\geq\abs{\bA_t}\} \right| \bA_t \right]  \right| \cH_t \right]\\
=& \EE \left[ \left. R(t, \bA_t) \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right]\\
\leq &\frac{1}{p^{\ast}} \EE \left[ \left. R(t, \bA_t) \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right]\\
\leq &\frac{1}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\abs{\bA_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right]\\
\leq &\frac{1}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\bO_t} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \right| \cH_t \right].
\end{align}
The third inequality is because $p_{\bA_t} \geq p^*$ and the forth inequality is by Lemma \ref{lem:DeltaEstimate}.
% $p_{\bA_t} \geq \frac{1}{n_0 + 1}$ by lemma \ref{lem:prefixRelation} and the result can be derived directly.

For the conjunctive case, by Lemma~\ref{lem:prefixexists} there exists a prefix $\bB_t$ of $\bA_t$ such that $p_{\wedge, t, \bB_t} \geq \frac{1}{2}f_{\wedge, t}^* - \gamma'_{|\bB_t|}$ and $R_{\wedge}(t, \bB_t) \geq \frac{1}{2} R_{\wedge}(t, \bA_t)$. 
By the assumption that $\gamma'_{|\bB_t|}\le \gamma'_K \le \frac{1}{4} f^*_{\wedge}
	\le \frac{1}{4} f^*_{\wedge,t}  $, we have
	$p_{\wedge, t, \bB_t} \geq \frac{1}{4}f_{\wedge, t}^*$.
Then we have
\begin{align}
&\EE_t[R_{\wedge}(t, \bA_t)] \leq 2 \EE[R_{\wedge}(t, \bB_t) \left. \right| \cH_t] \nonumber \\
=& 2 \EE \left[ \left. R_{\wedge}(t, \bB_t) \EE \left[ \left. \frac{1}{p_{\wedge, t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\} \right| \bB_t \right]  \right| \cH_t \right] \nonumber \\
=& 2\EE \left[ \left. R_{\wedge}(t, \bB_t) \frac{1}{p_{\wedge, t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right] \nonumber \\
\leq &\frac{8}{f_{\wedge, t}^{\ast}} \EE \left[ \left. R_{\wedge}(t, \bB_t) \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right]. \label{eq:btOneotbt}
% \leq &\frac{8}{f_{\wedge, t}^{\ast}} \EE \left[ \left. \sum_{k=1}^{\abs{\bB_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right]\\
% \leq &\frac{8}{f_{\wedge, t}^{\ast}} \EE \left[ \left. \sum_{k=1}^{\bO_t} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \right| \cH_t \right].
\end{align}
Next, we have
\begin{equation} \label{eq:andstarbt}
f_{\wedge}(A_t^*, \theta_*^{\top}x_t) \leq f_{\wedge}(A_t^*,\bU_t) \leq f_{\wedge}(\bA_t,\bU_t) \leq f_{\wedge}(\bB_t,\bU_t),
\end{equation}
where the first inequality is by Lemmas \ref{lem:estimateU} and \ref{lem:increasing},
	the second inequality is by the C$^3$-UCB algorithm, in which the best action $\bA_t$ is selected
	with respect to $\bU_t$, and the last inequality is by
	Lemma~\ref{lem:prefixRelation}~(2).

@@

\begin{lemma}
\label{lem:SumXiEstimateInDet}
Let $x_i \in \RR^{d \times 1}, 1 \leq i \leq n$. Then we have
$$
\det(I + \sum_{i=1}^n x_i x_i^{\top}) \geq 1 + \sum_{i=1}^n \norm{x_i}_2^2.
$$
\end{lemma}
\begin{proof}
Denote the eigenvalues of $I + \sum_{i=1}^n x_i x_i^{\top}$ by $1+\alpha_1,...,1+\alpha_d$ with $\alpha_j \geq 0$, $1\leq j\leq d$. Then
\begin{align*}
&\det(I + \sum_{i=1}^n x_i x_i^{\top})= \prod_{j=1}^d (1 + \alpha_j)\\
&\geq 1 +\sum_{j=1}^d \alpha_j =1-d + \sum_{i=1}^d (1+\alpha_i) \\
&=1-d + \trace(I + \sum_{i=1}^n x_i x_i^{\top})= 1-d + d + \sum_{i=1}^n \norm{x_i}_2^2\\
&=1 + \sum_{i=1}^n \norm{x_i}_2^2.
\end{align*}
\end{proof}

>>>>

To state our main theorem, we need some definitions. Let
$$
f_{\wedge}^* = \min_{1 \leq t \leq T} f_{\wedge, t}^{\ast}
$$
denote the best expected reward of all time under the conjunctive objective.
Let $p_{t, A}$ (resp., $p_{\wedge, t, A}$) be the {\em probability of full
	observation of $A$}, that is the probability that all base arms of $A = (a_1, \ldots, a_{\abs{A}})$ are observed under the disjunctive objective (resp., conjunctive objective) when under the context of time $t$, that is 
\begin{align*}
&p_{t, A} = \prod_{k=1}^{\abs{A}-1} (1 - \theta_{\ast}^{\top} x_{t, a_k^t}),\\
&p_{\wedge, t, A} = \prod_{k=1}^{\abs{A}-1} (\theta_{\ast}^{\top} x_{t, a_k^t}).
\end{align*}
We also define $p^* = \min_{1 \leq t \leq T} \min_{A \in \cS} ~ p_{t, A}$ to be the minimal probability over all time that an action could have all base arms observed under the disjunctive objective. The following is the main theorem on the regret achieved by our C$^3$-UCB algorithm, for both the disjunctive and conjunctive objectives.
\begin{theorem}
\label{thm:main}
Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_{K} \geq 0$ and 
$\gamma_k' = 1 - \gamma_k, \gamma'_K \leq \frac{1}{4}f_{\wedge}^{\ast}$.
Let regularization parameter $\lambda$ satisfy $\lambda \geq C_\gamma$, where $C_\gamma = \sum_{k=1}^{K} \gamma_k^2$. Then for any $\delta > 0$, with probability at least $1 - \delta$, the cumulative expected regrets of our algorithm, C$^3$-UCB for both the conjunctive and disjunctive objectives, satisfy
\begin{align}
R(T) &\le \frac{\sqrt{2}}{p^*} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))}  \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2}\right) } + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{p^*} \sqrt{TK} \ln \left(\frac{C_\gamma T}{\delta}\right) \right),
\end{align}
\begin{align}
R_{\wedge}(T) &\le \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} \nonumber \\
&\qquad \cdot \left(\sqrt{d\ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^2}\right) } + \sqrt{\lambda}\right) \nonumber \\
&=O\left(\frac{d}{f_{\wedge}^{\ast}}\sqrt{TK} \ln \left(\frac{C_\gamma T}{\delta}\right) \right).
\end{align}
%where
%\begin{align*}
%C_\gamma = \sum_{k=1}^{K} \gamma_k^2 = \sum_{k=1}^{K} (1 - \gamma_k')^2 \in [1, K].
%\end{align*}
\end{theorem}

@@

\begin{lemma}
\label{lem:DeltaEsimateWithP*}
Let $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$ and $\gamma_k' = 1 - \gamma_k$. 
%For any time $t$ and the chosen action $\bA_t = (\ba_1^t, \ldots, \ba_{\abs{A}}^t )$, we have $f(A_t^*, \bU_t) \leq f(A, \bU_t)$. 
Suppose the equation (\ref{eq:estimateTheta}) holds for time $t-1$. Then
$$
\EE_t[R(t, \bA_t)] \leq \frac{1}{p^*} \EE_t \left[ \bbeta_{t-1}(\delta) \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
$$
If we have
% $$
% \gamma_K \geq (1+\frac{1}{n_0})f^{\ast}, \qquad 
% $$
$$
\gamma_K' \le \frac{1}{4} f_{\wedge}^{\ast} = \frac{1}{4} \min_{1 \leq t \leq T} f_{\wedge, t}^{\ast},
$$
then
\begin{align*}
% &E[\Delta_{t, \bA_t}|\cH_t] \leq (n_0+1) \bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}},\\
&\EE_t [R_{\wedge}(t, \bA_t) ] \leq \frac{8}{f_{\wedge}^{\ast}} \EE_t \left[ \bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
\end{align*}
\end{lemma}
\begin{proof}
For the disjunctive case,  
\begin{align}
&\EE[R(t, \bA_t)|\cH_t]  \nonumber \\
=& \EE \left[ \left. R(t, \bA_t) \EE \left[ \left. \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t\geq\abs{\bA_t}\} \right| \bA_t \right]  \right| \cH_t \right]
	\label{eq:addOt}\\
=& \EE \left[ \left. R(t, \bA_t) \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right] \label{eq:removeE}\\
\leq &\frac{1}{p^{\ast}} \EE \left[ \left. R(t, \bA_t) \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right] \label{eq:relaxpstar}\\
\leq &\frac{1}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\abs{\bA_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right] \label{eq:useradiusbound}\\
\leq &\frac{1}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\bO_t} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \right| \cH_t \right].
\label{eq:relaxindicator}
\end{align}
Equality \eqref{eq:addOt} is because when $\bA_t$ is fixed, $p_{t, \bA_t}$
	is the probability of $\bO_t\geq\abs{\bA_t}$, and thus the inner expectation is $1$.
Equality \eqref{eq:removeE} is because when the history $\cH_t$ is fixed,
	$\bA_t$ is fixed by the deterministic algorithm C$^3$-UCB, and thus there is no
	need to write the conditional expectation.
Inequality~\eqref{eq:relaxpstar} is because $p_{t,\bA_t} \geq p^*$ by definition.
Inequality~\eqref{eq:useradiusbound} is by Lemma \ref{lem:DeltaEstimate} and
	the fact $f(A_t^*, \bU_t) \leq f(\bA_t, \bU_t)$, which is true because
	algorithm C$^3$-UCB selects action $\bA_t$ as the best action with respect to
	$\bU_t$.
Inequality~\eqref{eq:relaxindicator} is by simply arguing on the two cases for
	the indicator function $\bOne\{\bO_t\geq\abs{\bA_t}\}$.
% $p_{\bA_t} \geq \frac{1}{n_0 + 1}$ by lemma \ref{lem:prefixRelation} and the result can be derived directly.

For the conjunctive case, by Lemma~\ref{lem:prefixexists} there exists a prefix $\bB_t$ of $\bA_t$ such that $p_{\wedge, t, \bB_t} \geq \frac{1}{2}f_{\wedge, t}^* - \gamma'_{|\bB_t|}$ and $R_{\wedge}(t, \bB_t) \geq \frac{1}{2} R_{\wedge}(t, \bA_t)$. 
By the assumption that $\gamma'_{|\bB_t|}\le \gamma'_K \le \frac{1}{4} f^*_{\wedge}
	\le \frac{1}{4} f^*_{\wedge,t}  $, we have
	$p_{\wedge, t, \bB_t} \geq \frac{1}{4}f_{\wedge, t}^*$.
Then, similar to the derivation from Eq.\eqref{eq:addOt} to Eq.\eqref{eq:relaxpstar},
	we have
\begin{align}
&\EE_t[R_{\wedge}(t, \bA_t)] \leq 2 \EE[R_{\wedge}(t, \bB_t) \left. \right| \cH_t] \nonumber \\
=& 2 \EE \left[ \left. R_{\wedge}(t, \bB_t) \EE \left[ \left. \frac{1}{p_{\wedge, t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\} \right| \bB_t \right]  \right| \cH_t \right] \nonumber \\
=& 2\EE \left[ \left. R_{\wedge}(t, \bB_t) \frac{1}{p_{\wedge, t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right] \nonumber \\
\leq &\frac{8}{f_{\wedge, t}^{\ast}} \EE \left[ \left. R_{\wedge}(t, \bB_t) \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right]. \label{eq:btOneotbt}
% \leq &\frac{8}{f_{\wedge, t}^{\ast}} \EE \left[ \left. \sum_{k=1}^{\abs{\bB_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right]\\
% \leq &\frac{8}{f_{\wedge, t}^{\ast}} \EE \left[ \left. \sum_{k=1}^{\bO_t} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \right| \cH_t \right].
\end{align}
Next, we have
\begin{equation} \label{eq:andstarbt}
f_{\wedge}(A_t^*, \theta_*^{\top}x_t) \leq f_{\wedge}(A_t^*,\bU_t) \leq f_{\wedge}(\bA_t,\bU_t) \leq f_{\wedge}(\bB_t,\bU_t),
\end{equation}
where the first inequality is by Lemmas \ref{lem:estimateU} and \ref{lem:increasing},
	the second inequality is by the C$^3$-UCB algorithm, in which the best action $\bA_t$ is selected
	with respect to $\bU_t$, and the last inequality is by
	Lemma~\ref{lem:prefixRelation}~(2).

@@

\begin{lemma}
\label{lem:SumXiEstimateInDet}
Let $x_i \in \RR^{d \times 1}, 1 \leq i \leq n$. Then we have
$$
\det\left(I + \sum_{i=1}^n x_i x_i^{\top}\right) \geq 1 + \sum_{i=1}^n \norm{x_i}_2^2.
$$
\end{lemma}
\begin{proof}
Denote the eigenvalues of $I + \sum_{i=1}^n x_i x_i^{\top}$ by $1+\alpha_1,...,1+\alpha_d$ with $\alpha_j \geq 0$, $1\leq j\leq d$. Then
\begin{align*}
&\det(I + \sum_{i=1}^n x_i x_i^{\top})= \prod_{j=1}^d (1 + \alpha_j)\\
&\geq 1 +\sum_{j=1}^d \alpha_j =1-d + \sum_{i=1}^d (1+\alpha_i) \\
&=1-d + \trace(I + \sum_{i=1}^n x_i x_i^{\top})= 1-d + d + \sum_{i=1}^n \norm{x_i}_2^2\\
&=1 + \sum_{i=1}^n \norm{x_i}_2^2.
\end{align*}
\end{proof}
